MSGenova (Clean version) Rmakrdown Report

### 2022-09-13 script/ms_report.Rmd
update report, and change writing of config file, update report rule

---
title: "MSGenova (Clean version) Rmakrdown Report"
author: "Gabrièle Adam"
date: "2022-08-16"
output:
    html_document:
        highlight: tango
        number_sections: no
        theme: default
        toc: yes
        toc_depth: 3
        toc_float:
            collapsed: no
            smooth_scroll: yes
params:
  DP.min: DP.min
  AR.min: AR.min
  result_dir: result_dir
  rmd: "report.Rmd"
  sample: sample
  trimmo: trimmo
  multiqc: multiqc
  vector: vector
  flagstat: flagstat
  haplotype_caller: haplotype_caller_vcf
---

```{r setup, include=FALSE}
workspace<-params$result_dir
knitr::opts_chunk$set(echo = TRUE,root.dir=workspace)
```

```{r load_library, echo=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(DT)
library(ggplot2)
library(plyr)
library(dplyr)
library(yaml)
vector_calling<-params$vector
run_chromomap<-FALSE

DP.min<-params$DP.min
AR.min<-params$AR.min
#workspace<-params$result_dir

sheet<-params$sample
trimmo<- params$trimmo
flagstat<- params$flagstat
haplotype_caller<- params$haplotype_caller
#setwd(workspace)
```

@@

read_file<-function(pathway){
sample<-list.files(pathway)
vector<-list.files(paste0(pathway,"/",sample[1]))
i=0
for (spl in sample){
  for (v in vector){
    table<-read.table(paste0(pathway,"/",spl,"/",v,"/5.",v,"_insertion.annotated.bed"),header = FALSE,col.names = c("Chr","Breakpoint","SuppRead","TDNA_info","Orientation","Freq","Annotation"))
    if(! nrow(table)==0){
      #add sample and vector columns
      table["sample"]<-spl
      table["vector"]<- v
      if(i==0){
        final_table<-table
        i=i+1
      }else{
        final_table<-rbind(table,final_table)
      }
      i=i+1
    }
  }
}
return(final_table)
}
```

@@

```{r define_variable}
pathway=paste0(workspace,"/05_tdnascan/")
sample<-list.files(pathway)
print(sample)
vector<-list.files(paste0(pathway,"/",sample[1]))
print(vector)
# read sample file
sample_sheet<-read.table(paste0(workspace,"/genome/sample_sheet"),sep="\t",header=TRUE)
```

```{r read_input}
final_tab<-read_file(pathway)
DT::datatable(final_tab)
final_tab2<- final_tab %>% separate(col=TDNA_info,sep=",",into=c("tdna_st","tdna_end"))  %>% mutate(tdna_st=str_replace(tdna_st,"tdna_st:","")) %>% mutate(tdna_end=str_replace(tdna_end,"tdna_end:","")) %>% mutate(tdna_st=str_replace(tdna_st,"-","0")) %>% mutate(tdna_end=str_replace(tdna_end,"-","0")) %>% mutate(tdna_st=as.numeric(as.character(tdna_st))) %>% mutate(tdna_end=as.numeric(as.character(tdna_end))) %>% mutate(length_break=abs(tdna_end - tdna_st)) %>% mutate(breakpoint_end=Breakpoint+length_break) %>% mutate(name=paste(Chr,Breakpoint,vector,sample,sep="-"))  %>% select(name,Chr,Breakpoint,breakpoint_end,vector,sample)
DT::datatable(final_tab2)
```

## By vector, how many insertions are there by chromosome ?{.tabset .tabset-fade}

```{r barplot_snp, results='asis'}
#For one vector, look if it insert at the same position for several sample with a threshold for space between position
for (v in vector){
  sub_table<-filter(final_tab,vector==v) %>% group_by(Chr,sample) %>% dplyr::count(name="nbr_inster_by_chr")
  #datatable(sub_table)
  #count_tab<- count(sub_table,Chr)
  if(nrow(sub_table)> 0){
    cat("###",v,"\n")
  g<- ggplot(sub_table,aes(x=Chr,y=nbr_inster_by_chr,fill=sample))+geom_bar(stat="identity")+xlab("Numbre of insertion by chromosome")+ylab("Chromosome")+ggtitle(paste0("Number of in sertion of vector ",v," by chromosome"))+theme_bw()
  print(g)
  }
  cat('\n\n')
}
```

## What are the size between the insertion

```{r length_between_insertion}
list_of_size<- freq_length_bewteen_insertion(final_tab)
#size<-as.data.frame(list_of_size,rep("insertion",times=length(list_of_size)))
#ggplot(size,aes(x=list_of_size))+geom_density()+xlim(min(list_of_size),max(list_of_size))
size<-as.data.frame(count(list_of_size))
ggplot(size,aes(x=x,y=freq))+geom_point()+ylim(0,(max(size$freq)+1))
ggplot(size,aes(x=factor(x)))+geom_bar(stat = "count")
```

## Are there the same vector insertion for several sample - New version ?
```{r essai_shiny}
#list_size<-freq_length_bewteen_insertion(final_tab)
#cts<-count(list_size)
#print(cts)
#final_tab<-read_file("/home/gabriele.adam/Documents/Msgenova/sortie_msgenova/05_tdnascan/data")
essai<-search_identical_insertion(final_tab,5000)
DT::datatable(essai)
```

`r ifelse(vector_calling,"# how many samples have insertion of which vector and which do not have any - version barplot?","")`

```{r plot_by_chr_tuto, eval=TRUE, echo = TRUE}
#Establish list of sample without any insertion
no_insertion<-c()
#gff3<-"~/Documents/Msgenova/ref_arabido/Ath-Col0-tair10-WG/Arabidopsis_thaliana.TAIR10.41.gff3"
# read gff3 and extract size of chromosome ?? - > Jump to final table
##sequence-region   1 1 30427671
##sequence-region   2 1 19698289
##sequence-region   3 1 23459830
##sequence-region   4 1 18585056
##sequence-region   5 1 26975502
##sequence-region   Mt 1 366924
##sequence-region   Pt 1 154478
table_size_chr<-data.frame(c(paste0("Chr",c(1:5,"Mt","Pt"))),c(30427671,19698289,23459830,18585056,26975502,366924,154478))
colnames(table_size_chr)<-c("Chr","Size")
for (spl in unique(final_tab$sample)){
  sub_table<-dplyr::filter(final_tab,sample==spl)
  if(nrow(sub_table)==0){
    print("yes")
    no_insertion<-append(no_insertion,spl)
  }else{
 bp<- graphics::barplot(table_size_chr$Size,border=NA,names.arg = table_size_chr$Chr,main=paste0("Barplot showing plasmidic insertion on sample ",spl),horiz=TRUE)
 rownames(bp)<-table_size_chr$Chr
    with(sub_table,segments(sub_table$Breakpoint,bp[sub_table$Chr,]-0.5,sub_table$Breakpoint,bp[sub_table$Chr,]+0.5),bp[sub_table$Chr,]+0.5,col=sub_table$vector)
  #print(bp)
  }
}
print("List of vector without any plasmid insertion")
print(no_insertion)

@@

`r ifelse(run_chromomap,"#how many samples have insertion of which vector and which do not have any - version chromoMap","")`

```{r chromomap , eval =run_chromomap, echo=run_chromomap}
library(chromoMap)
path<-"/home/gabriele.adam/Documents/Msgenova/sortie_msgenova/chromomap/"
chr<-paste0(path,"chr_size")
annotation<-paste0(path,"annotation.tsv")
for (spl in unique(final_tab$sample)){
  sub_table<-dplyr::filter(final_tab,sample==spl)
  if(nrow(sub_table)==0){
    print("yes")
    no_insertion<-append(no_insertion,spl)
  }else{
 bp<- barplot(table_size_chr$Size,border=NA,names.arg = table_size_chr$Chr,main=paste0("Barplot showing plasmidic insertion on sample ",spl))
 rownames(bp)<-table_size_chr$Chr
  with(sub_table,segments(bp[sub_table$Chr,]-0.5,sub_table$Breakpoint,bp[sub_table$Chr,]+0.5),sub_table$Breakpoint,col=sub_table$vector)
  }}
chromoMap(chr,annotation)
sub_table<-dplyr::filter(final_tab,sample==spl) %>% separate(col=TDNA_info,sep=",",into=c("tdna_st","tdna_end"))  %>% mutate(tdna_st=str_replace(tdna_st,"tdna_st:","")) %>% mutate(tdna_end=str_replace(tdna_end,"tdna_end:","")) %>% mutate(tdna_st=str_replace(tdna_st,"-","0")) %>% mutate(tdna_end=str_replace(tdna_end,"-","0")) %>% mutate(tdna_st=as.numeric(as.character(tdna_st))) %>% mutate(tdna_end=as.numeric(as.character(tdna_end))) %>% mutate(length_break=abs(tdna_end - tdna_st)) %>% mutate(breakpoint_end=Breakpoint+length_break) %>% mutate(name=paste(Chr,Breakpoint,vector,sample,sep="-"))  %>% select(name,Chr,Breakpoint,breakpoint_end,vector)
write.table(sub_table,file=paste0(path,"annot2.tsv"),sep="\t",row.names = FALSE)
chromoMap(chr,paste0(path,"annot2.tsv"))
```

@@

```{r trimmomatic}
dat<-read.table(text=gsub(" ","\t",readLines(trimmo)),  header=TRUE)
DT::datatable(dat)
```

@@

```{r flagstat}
library(tidyverse)
file <- flagstat
if(file.exists(file)){
    flagstat <- read.table(file, sep="\t", header=FALSE) %>% 
      dplyr::mutate(mapped=V3-V4, unmapped=V2-V3, supplementary=V4, 
                    mapped.ratio=round(mapped/(mapped+unmapped)*100,2), 
                    unmapped.ratio=round(unmapped/(mapped+unmapped)*100,2),
                    supplementary.ratio=round(supplementary/(mapped+unmapped)*100,2)) %>% dplyr::select(-V2, -V3, -V4)
    names(flagstat) <- c("sample", "mapped", "unmapped", "supplementary", "mapped.ratio", "unmapped.ratio", "supplementary.ratio")

@@

### Coverage analysis
```{r coverage}
cov.list <- lapply(sample_sheet$sample, function(x){
  file <- paste0(workspace,"/03_mapping/coverage/", x, ".coverage")
  if(file.exists(file)){

@@

#### Depth analysis
```{r depth_analysis}
cov.list <- lapply(sample_sheet$sample, function(x){

  file <- paste0(workspace,"/03_mapping/depth/", x, ".depth")
  if(file.exists(file)){
    read.table(file, sep="\t")  %>% dplyr::mutate(plant = x)
  }
})

depth <- purrr::reduce(cov.list, rbind)
colnames(depth) <- c("rname", "position", "depth", "plant")
#depth <- depth %>% left_join(bed.tab, by="rname")  %>% filter(position >=start, position <= stop)
heading<-as.character(unique(sample_sheet$sample))
for (i in unique(cov$rname)){
 cat("####",i,"\n")
  p <- ggplot(data = filter(depth,rname== i)) + geom_line(aes(x=position, y=depth, color=plant))   + theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Depth of coverage")+facet_grid(~rname, scales = "free") 
  print(p)
cat('\n\n')
}

@@

```{r , echo=FALSE, warning=FALSE, error=FALSE}
vcfR.inst <- vcfR::read.vcfR(haplotype_caller)

>>>>

---
title: "MSGenova (Clean version) Rmakrdown Report"
author: "Gabrièle Adam"
date: "2022-08-16"
output:
    html_document:
        highlight: tango
        number_sections: no
        theme: default
        toc: yes
        toc_depth: 3
        toc_float:
            collapsed: no
            smooth_scroll: yes
params:
  DP.min: DP.min
  AR.min: AR.min
  result_dir: result_dir
  rmd: "report.Rmd"
---

```{r setup, include=FALSE}
workspace<-"/home/gabriele.adam/Documents/Msgenova/sorties_report/results"
knitr::opts_chunk$set(root.dir=workspace, echo = FALSE , warning=FALSE, error=FALSE, message=FALSE)
```

```{r load_library, echo=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(DT)
library(ggplot2)
library(plyr)
library(dplyr)
library(yaml)
vector_calling<-params$vector

DP.min<-5
AR.min<-0.2

sheet<-params$sample
trimmo<- params$trimmo
flagstat<- params$flagstat
haplotype_caller<- params$haplotype_caller

config_advanced_file<-paste0(workspace,"/00_logs/config_advanced")
config_ad<-read_yaml(config_advanced_file)
```

@@

read_file<-function(pathway){
sample<-list.files(pathway)
vector<-list.files(paste0(pathway,"/",sample[1]))
i=0
for (spl in sample){
  for (v in vector){
    table<-read.table(paste0(pathway,"/",spl,"/",v,"/5.",v,"_insertion.annotated.bed"),header = FALSE,col.names = c("Chr","Breakpoint","SuppRead","TDNA_info","Orientation","Freq","Annotation"))
    if(! nrow(table)==0){
      #add sample and vector columns
      table["sample"]<-spl
      table["vector"]<- v
      if(i==0){
        final_table<-table
        i=i+1
      }else{
        final_table<-rbind(table,final_table)
      }
      i=i+1
    }
  }
}
if (exists("final_table")){
return(final_table)}
else{return(table)}
}
```

@@

```{r define_variable}
pathway=paste0(workspace,"/05_tdnascan")
sample<-list.files(pathway)
print(sample)
vector<-list.files(paste0(pathway,"/",sample[1]))
print(vector)
# read sample file
sample_sheet<-read.table(paste0(workspace,"/genome/sample_sheet"),sep="\t",header=TRUE)
```

```{r read_input}
final_tab<-read_file(pathway)
DT::datatable(final_tab)
if(nrow(final_tab) >0){
final_tab2<- final_tab %>% separate(col=TDNA_info,sep=",",into=c("tdna_st","tdna_end"))  %>% mutate(tdna_st=str_replace(tdna_st,"tdna_st:","")) %>% mutate(tdna_end=str_replace(tdna_end,"tdna_end:","")) %>% mutate(tdna_st=str_replace(tdna_st,"-","0")) %>% mutate(tdna_end=str_replace(tdna_end,"-","0")) %>% mutate(tdna_st=as.numeric(as.character(tdna_st))) %>% mutate(tdna_end=as.numeric(as.character(tdna_end))) %>% mutate(length_break=abs(tdna_end - tdna_st)) %>% mutate(breakpoint_end=Breakpoint+length_break) %>% mutate(name=paste(Chr,Breakpoint,vector,sample,sep="-"))  %>% select(name,Chr,Breakpoint,breakpoint_end,vector,sample)
DT::datatable(final_tab2)
presence_of_insertion<-TRUE
}else{
  presence_of_insertion<-FALSE
}
```

`r ifelse(presence_of_insertion,"## By vector, how many insertions are there by chromosome ?{.tabset .tabset-fade}","")`

```{r barplot_snp, results='asis',eval=presence_of_insertion}
#For one vector, look if it insert at the same position for several sample with a threshold for space between position
for (v in vector){
  sub_table<-filter(final_tab,vector==v) %>% group_by(Chr,sample) %>% dplyr::count(name="nbr_inster_by_chr")
  #datatable(sub_table)
  #count_tab<- count(sub_table,Chr)
  if(nrow(sub_table)> 0){
    cat("###",v,"\n")
  g<- ggplot(sub_table,aes(x=Chr,y=nbr_inster_by_chr,fill=sample))+geom_bar(stat="identity",position=position_dodge())+ylab("Numbre of insertion by chromosome")+xlab("Chromosome")+ggtitle(paste0("Number of in sertion of vector ",v," by chromosome"))+theme_bw()
  print(g)
  }
  cat('\n\n')
}
```

# Does several samples have the same insertions ?
```{r same_insertions, eval=presence_of_insertion}
essai<-search_identical_insertion(final_tab,5000)
DT::datatable(essai)
```

`r ifelse(vector_calling,"# Barplot representation of the insertions in each sample","")`

```{r plot_by_chr_tuto, eval=presence_of_insertion, echo = presence_of_insertion}
#Establish list of sample without any insertion
no_insertion<-c()
#gff3<-"~/Documents/Msgenova/ref_arabido/Ath-Col0-tair10-WG/Arabidopsis_thaliana.TAIR10.41.gff3"
# read gff3 and extract size of chromosome ?? - > Jump to final table
##sequence-region   1 1 30427671
##sequence-region   2 1 19698289
##sequence-region   3 1 23459830
##sequence-region   4 1 18585056
##sequence-region   5 1 26975502
##sequence-region   Mt 1 366924
##sequence-region   Pt 1 154478
table_size_chr<-data.frame(c(paste0("Chr",c(1:5,"Mt","Pt"))),c(30427671,19698289,23459830,18585056,26975502,366924,154478))
colnames(table_size_chr)<-c("Chr","Size")
for (spl in unique(final_tab$sample)){
  sub_table<-dplyr::filter(final_tab,sample==spl)
  if(nrow(sub_table)==0){
    print("yes")
    no_insertion<-append(no_insertion,spl)
  }else{
 bp<- graphics::barplot(table_size_chr$Size,border=NA,names.arg = table_size_chr$Chr,main=paste0("Barplot showing plasmidic insertion on sample ",spl),horiz=TRUE)
 rownames(bp)<-table_size_chr$Chr
    with(sub_table,segments(sub_table$Breakpoint,bp[sub_table$Chr,]-0.5,sub_table$Breakpoint,bp[sub_table$Chr,]+0.5),bp[sub_table$Chr,]+0.5,col=sub_table$vector)
  #print(bp)
  }
}
```

```{r no_insertions, eval=presence_of_insertion}
print("List of vector without any plasmid insertion")
print(no_insertion)

@@

@@

```{r trimmomatic}
trimmo<-paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
dat<-read.table(text=gsub(" ","\t",readLines(trimmo)),  header=TRUE)
DT::datatable(dat)
```

@@

```{r flagstat}
library(tidyverse)
file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
if(file.exists(file)){
    flagstat <- read.table(file, sep="\t", header=FALSE) %>% 
      dplyr::mutate(mapped=V3-V4, unmapped=V2-V3, supplementary=V4, 
                    mapped.ratio=round(mapped/(mapped+unmapped)*100,2), 
                    unmapped.ratio=round(unmapped/(mapped+unmapped)*100,2),
                    supplementary.ratio=round(supplementary/(mapped+unmapped)*100,2)) %>% dplyr::select(-V2, -V3, -V4)
    names(flagstat) <- c("sample", "mapped", "unmapped", "supplementary", "mapped.ratio", "unmapped.ratio", "supplementary.ratio")

@@

### Coverage analysis
```{r coverage}
cov.list <- lapply(sample_sheet$sample, function(x){
  file <- paste0(workspace,"/02_mapping/coverage/", x, ".coverage")
  if(file.exists(file)){

@@

### Depth analysis
```{r depth_analysis}
cov.list <- lapply(sample_sheet$sample, function(x){

  file <- paste0(workspace,"/02_mapping/depth/", x, ".depth")
  if(file.exists(file)){
    read.table(file, sep="\t")  %>% dplyr::mutate(plant = x)
  }
})

depth <- purrr::reduce(cov.list, rbind)
colnames(depth) <- c("rname", "position", "depth", "sample")
#depth <- depth %>% left_join(bed.tab, by="rname")  %>% filter(position >=start, position <= stop)
heading<-as.character(unique(sample_sheet$sample))
for (i in unique(cov$rname)){
 cat("####",i,"\n")
  p <- ggplot(data = filter(depth,rname== i)) + geom_line(aes(x=position, y=depth, color=sample))   + theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Depth of coverage")+facet_grid(~rname, scales = "free") 
  print(p)
cat('\n\n')
}

@@

```{r , echo=FALSE, warning=FALSE, error=FALSE}
haplotype_caller<-paste0(workspace,"/03_snv_indels_calling/",config_ad$project_name,"_snv_indel.vcf")
vcfR.inst <- vcfR::read.vcfR(haplotype_caller)
